---
title: "Analysis"
author: "Maryam Sadeghi"
date: "30/07/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Analysis of Data
The first step is to read the data 

```{r}
activity<-read.csv("activity.csv")
```
In the next step, we want to find the total number of steps taken per day by taking a histogram of the values.
```{r}
activity$day <-weekdays(as.Date(activity$date))
activity$DateTime<- as.POSIXct(activity$date, format="%Y-%m-%d")
clean <- activity[!is.na(activity$steps),]
sumTable <- aggregate(activity$steps ~ activity$date, FUN=sum, )
colnames(sumTable)<- c("Date", "Steps")
hist(sumTable$Steps, breaks=5, xlab="Steps", main = "Total Steps per Day")
```
New, let us find the mean and median of the total number of steps taken per day.
```{r}
mean(sumTable$Steps)
median(sumTable$Steps)
```
According to the above code, the mean and median are 10766.19 and 10765, respectively. The following plot indicates the average number of steps per day.
```{r}
intervalTable <- ddply(clean, .(interval), summarize, Avg = mean(steps))
p <- ggplot(intervalTable, aes(x=interval, y=Avg), xlab = "Interval", ylab="Average Number of Steps")
p + geom_line()+xlab("Interval")+ylab("Average Number of Steps")+ggtitle("Average Number of Steps per Interval")
```
Based on this plot, the highst point is approximately 220 and the lowest point is 0.
```{r}
maxSteps <- max(intervalTable$Avg)
intervalTable[intervalTable$Avg==maxSteps,1]
nrow(activity[is.na(activity$steps),])
```
Hence, the total number of missing values in the dataset is 2304.
Differences in activity patterns: Weekdays vs Weekends.
```{r}
avgTable <- ddply(clean, .(interval, day), summarize, Avg = mean(steps))
nadata<- activity[is.na(activity$steps),]
newdata<-merge(nadata, avgTable, by=c("interval", "day"))
newdata2<- newdata[,c(6,4,1,2,5)]
colnames(newdata2)<- c("steps", "date", "interval", "day", "DateTime")
mergeData <- rbind(clean, newdata2)
sumTable2 <- aggregate(mergeData$steps ~ mergeData$date, FUN=sum, )
colnames(sumTable2)<- c("Date", "Steps")
mean(sumTable2$Steps)
median(sumTable2$Steps)
```



